\pkg{cv}: An \proglang{R} Package for Cross-Validation of Regression Models

### 2024-01-04 scratch/notes_john/JSS-article/JSS-article-reduced.Rmd
combine, delete, and resize graphs

```{r mpg-horsepower-scatterplot}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "`mpg` vs `horsepower` for the `Auto` data"
plot(mpg ~ horsepower, data=Auto)
```
The relationship between the two variables is monotone, decreasing, and nonlinear. Following @JamesEtAl:2021, we'll consider approximating the relationship by a polynomial regression, with the degree of the polynomial $p$ ranging from 1 (a linear regression) to 10.[^log-trans] Polynomial fits  for $p = 1$ to $5$ are shown in the following figure:

@@

```{r mpg-horsepower-scatterplot-polynomials}
#| out.width = "100%",
#| fig.height = 5,
#| fig.cap = "`mpg` vs `horsepower` for the `Auto` data"
plot(mpg ~ horsepower, data=Auto)
horsepower <- with(Auto, 
                   seq(min(horsepower), max(horsepower), 
                       length=1000))
for (p in 1:5){
  m <- lm(mpg ~ poly(horsepower,p), data=Auto)
  mpg <- predict(m, newdata=data.frame(horsepower=horsepower))
  lines(horsepower, mpg, col=p + 1, lty=p, lwd=2)
}
legend("topright", legend=1:5, col=2:6, lty=1:5, lwd=2,
       title="Degree", inset=0.02)
```
The linear fit is clearly inappropriate; the fits for $p = 2$ (quadratic) through $4$ are very similar; and the fit for $p = 5$ may over-fit the data by chasing one or two relatively high `mpg` values at the right (but see the CV results reported below).

@@

```{r mpg-horsepower-MSE-se}
#| out.width = "100%",
#| fig.height = 5,
#| fig.cap = "Estimated squared error as a function of polynomial degree, $p$"
library("cv") # for mse() and other functions

@@

Finally, we extract and graph the adjusted MSEs for $10$-fold CV and the MSEs for LOO CV:
```{r polynomial-regression-CV-graph}
#| out.width = "100%",
#| fig.height = 5,
#| fig.cap = "Cross-validated 10-fold and LOO MSE as a function of polynomial degree, $p$"
cv.mse.10 <- sapply(cv.auto.10, function(x) x[["adj CV crit"]])
cv.mse.loo <- sapply(cv.auto.loo, function(x) x[["CV crit"]])
plot(c(1, 10), range(cv.mse.10, cv.mse.loo), type="n",
     xlab="Degree of polynomial, p",
     ylab="Cross-Validated MSE")
lines(1:10, cv.mse.10, lwd=2, lty=1, col=2, pch=16, type="b")
lines(1:10, cv.mse.loo, lwd=2, lty=2, col=3, pch=17, type="b")
legend("topright", inset=0.02,
       legend=c("10-Fold CV", "LOO CV"),
       lwd=2, lty=2:1, col=3:2, pch=17:16)
```

@@

```{r cross-validation-clusters}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "10-fold cluster-based cross-validation comparing random intercept models with varying fixed effects. The error bars show the 95% confidence interval around the CV estimate of the MSE for each model."
modlist <- models("~ 1"=mod.0, "~ 1 + x"=mod.1, 
                  "~ 1 + x + xm"=mod.2, "~ 1 + I(x - xm)"=mod.3)
cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)
plot(cvs_clusters, main="Model Comparison, Cluster-Based CV")
```

```{r cross-validation-cases}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "10-fold case-based cross-validation comparing random intercept models with varying fixed effects."
cvs_cases <- cv(modlist, data=Data, seed=9693)
plot(cvs_cases, main="Model Comparison, Case-Based CV")
```

@@

@@

A scatterplot matrix (using the \code{scatterplotMatrix()} function in the \pkg{car} package) of the numeric variables in the data reveals that the distributions of \code{income} and \code{women} are positively skewed, and that some of the relationships among the three predictors, and between the predictors and the response (i.e., \code{prestige}), are nonlinear:
```{r scatterplot-matrix}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Scatterplot matrix for the `Prestige` data."
library("car")
scatterplotMatrix(~ prestige + income + education + women,
                  data=Prestige, smooth=list(spread=FALSE))
```
The \code{powerTransform()} function in the \pkg{car} package transforms variables towards multivariate normality by a generalization of Box and Cox's maximum-likelihood-like approach [@BoxCox:1964]. Several "families" of power transformations can be used, including the original Box-Cox family, simple powers (and roots), and two adaptations of the Box-Cox family to data that may include negative values and zeros: the Box-Cox-with-negatives family and the Yeo-Johnson family; see @Weisberg:2014 [Chap. 8], and @FoxWeisberg:2019 [Chap. 3] for details. Because \code{women} has some 0 values, we use the Yeo-Johnson family:
```{r power-transform-Prestige}
trans <- powerTransform( cbind(income, education, women) ~ 1,
                         data=Prestige, family="yjPower")
summary(trans)
```
We thus have evidence of the desirability of transforming \code{income} (by the $1/3$ power) and \code{women} (by the $0.16$ power---which is close to the "0" power, i.e., the log transformation), but not \code{education}. Applying the "rounded" power transformations makes the predictors better-behaved:
```{r transformed-predictors}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Scatterplot matrix for the `Prestige` data with the predictors transformed."
P <- Prestige[, c("prestige", "income", "education", "women")]
(lambdas <- trans$roundlam)
names(lambdas) <- c("income", "education", "women")
for (var in c("income", "education", "women")){
  P[, var] <- yjPower(P[, var], lambda=lambdas[var])
}
summary(P)

scatterplotMatrix(~ prestige + income + education + women,
                  data=P, smooth=list(spread=FALSE))
```
Comparing the MSE for the regressions with the original and transformed predictors shows a advantage to the latter:
```{r prestige-regressions}
m.pres <- lm(prestige ~ income + education + women, data=Prestige)
m.pres.trans <- lm(prestige ~ income + education + women, data=P)
mse(Prestige$prestige, fitted(m.pres))
mse(P$prestige, fitted(m.pres.trans))
```
Similarly, component+residual plots for the two regressions, produced by the \code{crPlots()} function in the \pkg{car} package, suggest that the partial relationship of \code{prestige} to \code{income} is more nearly linear in the transformed data, but the transformation of \code{women] fails to capture what appears to be a slight quadratic partial relationship; the partial relationship of \code{prestige} to \code{education} is close to linear in both regressions:
```{r CR-plots-untransformed}
#| fig.cap = "Component+residual plots for the `Prestige` regression with the original predictors."
crPlots(m.pres)
```
```{r CR-plots-transformed}
#| fig.cap = "Component+residual plots for the `Prestige` regression with transformed predictors."
crPlots(m.pres.trans)
```

@@

To illustrate this process, we return to the \code{Auto} data set:
```{r Auto-redux}
summary(Auto)
xtabs(~ year, data=Auto)
xtabs(~ origin, data=Auto)
xtabs(~ cylinders, data=Auto)
```
We previously used the \code{Auto} here in a preliminary example where we employed CV to inform the selection of the order of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set:
```{r Auto-explore}
#| out.width = "100%",
#| fig.height = 7,
#| fig.cap = "Scatterplot matrix for the numeric variables in the `Auto` data"
Auto$cylinders <- factor(Auto$cylinders,
                         labels=c("3.4", "3.4", "5.6", "5.6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels=c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique=TRUE)
Auto$name <- NULL

@@

We begin with a "working model" that specifies linear partial relationships of the response to the numeric predictors:
```{r Auto-working-model}
#| out.width = "100%",
#| fig.height = 7,
#| fig.cap = "Component+residual plots for the working model fit to the `Auto` data"
m.auto <- lm(mpg ~ ., data = Auto)
summary(m.auto)

@@

The transformed numeric variables are much better-behaved:
```{r Auto-transformed-scatterplot-matrix}
#| out.width = "100%",
#| fig.height = 7,
#| fig.cap = "Scatterplot matrix for the transformed numeric variables in the `Auto` data"
scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
                  + acceleration, 
                  smooth=list(spread=FALSE), data=A)
```
And the partial relationships in the model fit to the transformed data are much more nearly linear:
```{r Auto-CR-plots-transformed}
#| out.width = "100%",
#| fig.height = 7,
#| fig.cap = "Component+residual plots for the model fit to the transformed `Auto` data"
crPlots(m)
```

>>>>

The relationship between the two variables is monotone, decreasing, and nonlinear. Following @JamesEtAl:2021, we'll consider approximating the relationship by a polynomial regression, with the degree of the polynomial $p$ ranging from 1 (a linear regression) to 10.[^log-trans] Polynomial fits  for $p = 1$ to $5$ are shown in the following figure:

@@

```{r mpg-horsepower-scatterplot-polynomials}
#| out.width = "50%",
#| fig.height = 5,
#| fig.cap = "`mpg` vs `horsepower` for the `Auto` data"
plot(mpg ~ horsepower, data=Auto)
horsepower <- with(Auto, 
                   seq(min(horsepower), max(horsepower), 
                       length=1000))
for (p in 1:5){
  m <- lm(mpg ~ poly(horsepower,p), data=Auto)
  mpg <- predict(m, newdata=data.frame(horsepower=horsepower))
  lines(horsepower, mpg, col=p + 1, lty=p, lwd=2)
}
legend("topright", legend=1:5, col=2:6, lty=1:5, lwd=2,
       title="Degree", inset=0.02)
```
The linear fit is clearly inappropriate; the fits for $p = 2$ (quadratic) through $4$ are very similar; and the fit for $p = 5$ may over-fit the data by chasing one or two relatively high `mpg` values at the right (but see the CV results reported below).

@@

```{r mpg-horsepower-MSE-se}
#| out.width = "50%",
#| fig.height = 5,
#| fig.cap = "Estimated squared error as a function of polynomial degree, $p$"
library("cv") # for mse() and other functions

@@

@@

```{r cross-validation-clusters}
#| fig.height = 5,
#| fig.subcap=rep("", 2),
#| fig.ncol = 2, out.width = "45%", fig.align = "center",
#| fig.cap = "10-fold (a) cluster-based and (b) case-based cross-validation comparing random intercept models with varying fixed effects. The error bars show the 95% confidence interval around the CV estimate of the MSE for each model."
modlist <- models("~ 1"=mod.0, "~ 1 + x"=mod.1, 
                  "~ 1 + x + xm"=mod.2, "~ 1 + I(x - xm)"=mod.3)

cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)
plot(cvs_clusters, main="Model Comparison, Cluster-Based CV")

cvs_cases <- cv(modlist, data=Data, seed=9693)
plot(cvs_cases, main="Model Comparison, Case-Based CV")
```

@@

# Cross-validating model selection

@@

A scatterplot matrix (using the \code{scatterplotMatrix()} function in the \pkg{car} package) of the numeric variables in the data reveals that the distributions of \code{income} and \code{women} are positively skewed, and that some of the relationships among the three predictors, and between the predictors and the response (i.e., \code{prestige}), are nonlinear:
```{r scatterplot-matrix}
#| fig.height = 5,
#| fig.cap = "Scatterplot matrix for the `Prestige` data, (a) untransformed and (b) transformed",
#| fig.subcap=rep("", 2),
#| fig.ncol = 2, out.width = "50%", fig.align = "center"
library("car")
scatterplotMatrix(~ prestige + income + education + women,
                  data=Prestige, smooth=list(spread=FALSE))

trans <- powerTransform( cbind(income, education, women) ~ 1,
                         data=Prestige, family="yjPower")
summary(trans)

P <- Prestige[, c("prestige", "income", "education", "women")]
(lambdas <- trans$roundlam)
names(lambdas) <- c("income", "education", "women")
for (var in c("income", "education", "women")){
  P[, var] <- yjPower(P[, var], lambda=lambdas[var])
}
scatterplotMatrix(~ prestige + income + education + women,
                  data=P, smooth=list(spread=FALSE))

```
The \code{powerTransform()} function in the \pkg{car} package transforms variables towards multivariate normality by a generalization of Box and Cox's maximum-likelihood-like approach [@BoxCox:1964]. Several "families" of power transformations can be used, including the original Box-Cox family, simple powers (and roots), and two adaptations of the Box-Cox family to data that may include negative values and zeros: the Box-Cox-with-negatives family and the Yeo-Johnson family; see @Weisberg:2014 [Chap. 8], and @FoxWeisberg:2019 [Chap. 3] for details. Because \code{women} has some 0 values, we use the Yeo-Johnson family:

We thus have evidence of the desirability of transforming \code{income} (by the $1/3$ power) and \code{women} (by the $0.16$ power---which is close to the "0" power, i.e., the log transformation), but not \code{education}. Applying the "rounded" power transformations makes the predictors better-behaved:

Comparing the MSE for the regressions with the original and transformed predictors shows a advantage to the latter:
```{r prestige-regressions}
m.pres <- lm(prestige ~ income + education + women, data=Prestige)
m.pres.trans <- lm(prestige ~ income + education + women, data=P)
mse(Prestige$prestige, fitted(m.pres))
mse(P$prestige, fitted(m.pres.trans))
```
Similarly, component+residual plots for the two regressions, produced by the \code{crPlots()} function in the \pkg{car} package, suggest that the partial relationship of \code{prestige} to \code{income} is more nearly linear in the transformed data, but the transformation of \code{women] fails to capture what appears to be a slight quadratic partial relationship; the partial relationship of \code{prestige} to \code{education} is close to linear in both regressions:

```{r CR-plots}
#| fig.height = 5,
#| fig.cap = "Component+residual plots for the `Prestige` regression with (a) the original predictors and (b) transformed predictors",
#| fig.subcap=rep("", 2),
#| fig.ncol = 1, out.width = "60%", fig.align = "center"
crPlots(m.pres)

crPlots(m.pres.trans)
```

@@

To illustrate this process, we return to the \code{Auto} data set:
```{r Auto-redux}
summary(Auto)
xtabs(~ year, data=Auto)
xtabs(~ origin, data=Auto)
xtabs(~ cylinders, data=Auto)
```
We previously used the \code{Auto} here in a preliminary example where we employed CV to inform the selection of the order of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set:
```{r Auto-explore}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Scatterplot matrix for the numeric variables in the `Auto` data"
Auto$cylinders <- factor(Auto$cylinders,
                         labels=c("3.4", "3.4", "5.6", "5.6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels=c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique=TRUE)
Auto$name <- NULL

@@

We begin with a "working model" that specifies linear partial relationships of the response to the numeric predictors:
```{r Auto-working-model}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Component+residual plots for the working model fit to the `Auto` data"
m.auto <- lm(mpg ~ ., data = Auto)
summary(m.auto)

@@

The transformed numeric variables are much better-behaved:
```{r Auto-transformed-scatterplot-matrix}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Scatterplot matrix for the transformed numeric variables in the `Auto` data"
scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
                  + acceleration, 
                  smooth=list(spread=FALSE), data=A)
```
And the partial relationships in the model fit to the transformed data are much more nearly linear:
```{r Auto-CR-plots-transformed}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Component+residual plots for the model fit to the transformed `Auto` data"
crPlots(m)
```

### 2024-01-04 scratch/notes_john/JSS-article/JSS-article-reduced.tex
combine, delete, and resize graphs

\begin{CodeChunk}
\begin{CodeInput}
R> plot(mpg ~ horsepower, data=Auto)
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-scatterplot-1}

\caption[`mpg` vs `horsepower` for the `Auto` data]{`mpg` vs `horsepower` for the `Auto` data}\label{fig:mpg-horsepower-scatterplot}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-scatterplot-polynomials-1}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-MSE-se-1}

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 866709
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.27395
bias-adjusted cross-validation criterion = 19.25863
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

Finally, we extract and graph the adjusted MSEs for \(10\)-fold CV and
the MSEs for LOO CV:

\begin{CodeChunk}
\begin{CodeInput}
R> cv.mse.10 <- sapply(cv.auto.10, function(x) x[["adj CV crit"]])
R> cv.mse.loo <- sapply(cv.auto.loo, function(x) x[["CV crit"]])
R> plot(c(1, 10), range(cv.mse.10, cv.mse.loo), type="n",
+      xlab="Degree of polynomial, p",
+      ylab="Cross-Validated MSE")
R> lines(1:10, cv.mse.10, lwd=2, lty=1, col=2, pch=16, type="b")
R> lines(1:10, cv.mse.loo, lwd=2, lty=2, col=3, pch=17, type="b")
R> legend("topright", inset=0.02,
+        legend=c("10-Fold CV", "LOO CV"),
+        lwd=2, lty=2:1, col=3:2, pch=17:16)
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/polynomial-regression-CV-graph-1}

\caption[Cross-validated 10-fold and LOO MSE as a function of polynomial degree, $p$]{Cross-validated 10-fold and LOO MSE as a function of polynomial degree, $p$}\label{fig:polynomial-regression-CV-graph}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> modlist <- models("~ 1"=mod.0, "~ 1 + x"=mod.1, 
+                   "~ 1 + x + xm"=mod.2, "~ 1 + I(x - xm)"=mod.3)
R> cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)
R> plot(cvs_clusters, main="Model Comparison, Cluster-Based CV")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-clusters-1}

@@

\caption[10-fold cluster-based cross-validation comparing random intercept models with varying fixed effects]{10-fold cluster-based cross-validation comparing random intercept models with varying fixed effects. The error bars show the 95\% confidence interval around the CV estimate of the MSE for each model.}\label{fig:cross-validation-clusters}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> cvs_cases <- cv(modlist, data=Data, seed=9693)
R> plot(cvs_cases, main="Model Comparison, Case-Based CV")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-cases-1}

\caption[10-fold case-based cross-validation comparing random intercept models with varying fixed effects]{10-fold case-based cross-validation comparing random intercept models with varying fixed effects.}\label{fig:cross-validation-cases}
\end{figure}
\end{CodeChunk}

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("car")
\end{CodeInput}
\begin{CodeOutput}
Loading required package: carData
\end{CodeOutput}
\begin{CodeInput}
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=Prestige, smooth=list(spread=FALSE))
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/scatterplot-matrix-1}

@@

\caption[Scatterplot matrix for the `Prestige` data]{Scatterplot matrix for the `Prestige` data.}\label{fig:scatterplot-matrix}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> trans <- powerTransform( cbind(income, education, women) ~ 1,
+                          data=Prestige, family="yjPower")
R> summary(trans)
\end{CodeInput}
\begin{CodeOutput}
yjPower Transformations to Multinormality 
          Est Power Rounded Pwr Wald Lwr Bnd Wald Upr Bnd
income       0.2678        0.33       0.1051       0.4304
education    0.5162        1.00      -0.2822       1.3145
women        0.1630        0.16       0.0112       0.3149

 Likelihood ratio test that all transformation parameters are equal to 0
                               LRT df      pval
LR test, lambda = (0 0 0) 15.73879  3 0.0012827
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> P <- Prestige[, c("prestige", "income", "education", "women")]
R> (lambdas <- trans$roundlam)
\end{CodeInput}
\begin{CodeOutput}
   income education     women 
0.3300000 1.0000000 0.1630182 
\end{CodeOutput}
\begin{CodeInput}
R> names(lambdas) <- c("income", "education", "women")
R> for (var in c("income", "education", "women")){
+   P[, var] <- yjPower(P[, var], lambda=lambdas[var])
+ }
R> summary(P)
\end{CodeInput}
\begin{CodeOutput}
    prestige         income        education          women      
 Min.   :14.80   Min.   :22.15   Min.   : 6.380   Min.   :0.000  
 1st Qu.:35.23   1st Qu.:44.17   1st Qu.: 8.445   1st Qu.:1.731  
 Median :43.60   Median :50.26   Median :10.540   Median :3.362  
 Mean   :46.83   Mean   :50.76   Mean   :10.738   Mean   :3.502  
 3rd Qu.:59.27   3rd Qu.:56.24   3rd Qu.:12.648   3rd Qu.:5.591  
 Max.   :87.20   Max.   :83.62   Max.   :15.970   Max.   :6.830  
\end{CodeOutput}
\begin{CodeInput}
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=P, smooth=list(spread=FALSE))
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/transformed-predictors-1}

\caption[Scatterplot matrix for the `Prestige` data with the predictors transformed]{Scatterplot matrix for the `Prestige` data with the predictors transformed.}\label{fig:transformed-predictors}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m.pres)
\end{CodeInput}
\begin{figure}

{\centering \includegraphics{JSS-article-reduced_files/figure-latex/CR-plots-untransformed-1}

@@

\caption[Component+residual plots for the `Prestige` regression with the original predictors]{Component+residual plots for the `Prestige` regression with the original predictors.}\label{fig:CR-plots-untransformed}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m.pres.trans)
\end{CodeInput}
\begin{figure}

{\centering \includegraphics{JSS-article-reduced_files/figure-latex/CR-plots-transformed-1}

\caption[Component+residual plots for the `Prestige` regression with transformed predictors]{Component+residual plots for the `Prestige` regression with transformed predictors.}\label{fig:CR-plots-transformed}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/Auto-explore-1}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/Auto-working-model-1}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/Auto-transformed-scatterplot-matrix-1}

@@

{\centering \includegraphics[width=1\linewidth]{JSS-article-reduced_files/figure-latex/Auto-CR-plots-transformed-1}

>>>>

@@

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-scatterplot-polynomials-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-MSE-se-1}

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 577350
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.19107
bias-adjusted cross-validation criterion = 19.18016
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> modlist <- models("~ 1"=mod.0, "~ 1 + x"=mod.1, 
+                   "~ 1 + x + xm"=mod.2, "~ 1 + I(x - xm)"=mod.3)
R> 
R> cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)
R> plot(cvs_clusters, main="Model Comparison, Cluster-Based CV")
R> 
R> cvs_cases <- cv(modlist, data=Data, seed=9693)
R> plot(cvs_cases, main="Model Comparison, Case-Based CV")
\end{CodeInput}
\begin{figure}

{\centering \subfloat[\label{fig:cross-validation-clusters-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-clusters-1} }\subfloat[\label{fig:cross-validation-clusters-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-clusters-2} }

@@

\caption[10-fold (a) cluster-based and (b) case-based cross-validation comparing random intercept models with varying fixed effects]{10-fold (a) cluster-based and (b) case-based cross-validation comparing random intercept models with varying fixed effects. The error bars show the 95\% confidence interval around the CV estimate of the MSE for each model.}\label{fig:cross-validation-clusters}
\end{figure}
\end{CodeChunk}

@@

\hypertarget{cross-validating-model-selection}{%
\section{Cross-validating model
selection}\label{cross-validating-model-selection}}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("car")
\end{CodeInput}
\begin{CodeOutput}
Loading required package: carData
\end{CodeOutput}
\begin{CodeInput}
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=Prestige, smooth=list(spread=FALSE))
R> 
R> trans <- powerTransform( cbind(income, education, women) ~ 1,
+                          data=Prestige, family="yjPower")
R> summary(trans)
\end{CodeInput}
\begin{CodeOutput}
yjPower Transformations to Multinormality 
          Est Power Rounded Pwr Wald Lwr Bnd Wald Upr Bnd
income       0.2678        0.33       0.1051       0.4304
education    0.5162        1.00      -0.2822       1.3145
women        0.1630        0.16       0.0112       0.3149

 Likelihood ratio test that all transformation parameters are equal to 0
                               LRT df      pval
LR test, lambda = (0 0 0) 15.73879  3 0.0012827
\end{CodeOutput}
\begin{CodeInput}
R> P <- Prestige[, c("prestige", "income", "education", "women")]
R> (lambdas <- trans$roundlam)
\end{CodeInput}
\begin{CodeOutput}
   income education     women 
0.3300000 1.0000000 0.1630182 
\end{CodeOutput}
\begin{CodeInput}
R> names(lambdas) <- c("income", "education", "women")
R> for (var in c("income", "education", "women")){
+   P[, var] <- yjPower(P[, var], lambda=lambdas[var])
+ }
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=P, smooth=list(spread=FALSE))
\end{CodeInput}
\begin{figure}

{\centering \subfloat[\label{fig:scatterplot-matrix-1}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/scatterplot-matrix-1} }\subfloat[\label{fig:scatterplot-matrix-2}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/scatterplot-matrix-2} }

@@

\caption[Scatterplot matrix for the `Prestige` data, (a) untransformed and (b) transformed]{Scatterplot matrix for the `Prestige` data, (a) untransformed and (b) transformed}\label{fig:scatterplot-matrix}
\end{figure}
\end{CodeChunk}

@@

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m.pres)
R> 
R> crPlots(m.pres.trans)
\end{CodeInput}
\begin{figure}

{\centering \subfloat[\label{fig:CR-plots-1}]{\includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/CR-plots-1} }\newline\subfloat[\label{fig:CR-plots-2}]{\includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/CR-plots-2} }\newline

@@

\caption[Component+residual plots for the `Prestige` regression with (a) the original predictors and (b) transformed predictors]{Component+residual plots for the `Prestige` regression with (a) the original predictors and (b) transformed predictors}\label{fig:CR-plots}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/Auto-explore-1}

@@

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/Auto-working-model-1}

@@

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/Auto-transformed-scatterplot-matrix-1}

@@

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/Auto-CR-plots-transformed-1}
